---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
```{r}
dbGetQuery(sc, "DESC orders_tbl")

day_order_txt="
SELECT order_dow
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_dow
ORDER BY n_orders DESC
LIMIT 50"
day_order <-
  dbGetQuery(sc, day_order_txt)
day_order
```

* The most popular hour of doing order? 
```{r}
hour_order_txt="
SELECT order_hour_of_day
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY n_orders DESC
LIMIT 50"
hour_order <-
  dbGetQuery(sc,hour_order_txt)
hour_order
```

* The most reordered products?
```{r}
products_reorded_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   sum(reordered) AS n_reorders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_reorders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_reorded <-
  dbGetQuery(sc, products_reorded_txt)

products_reorded
```

* The less reorderd product?

```{r}
products_reorded_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   sum(reordered) AS n_reorders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_reorders
    ) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
where n_reorders!=0
limit 30
"

products_reorded_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   sum(reordered) AS n_reorders
    FROM order_products__prior_tbl
    where reordered!=0
    GROUP BY product_id
    ORDER BY n_reorders
    limit 30
    ) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_reorded <-
  dbGetQuery(sc, products_reorded_txt)

products_reorded
```

* Plot histogram of number of products in an order.
```{r}
ord_prod_txt="SELECT order_id,  max(add_to_cart_order) as n_prod
    FROM order_products_train_tbl 
    group by order_id
    order by n_prod desc
    LIMIT 10"
ord_prod_txt="SELECT order_id,  count(1) as n_prod
    FROM order_products_train_tbl 
    group by order_id
    order by n_prod desc
"
ord_prod <-
  dbGetQuery(sc, ord_prod_txt)
ord_prod

dbGetQuery(sc, "SELECT order_id,  count(1) as n_prod
    FROM order_products_train_tbl 
    group by order_id
    order by n_prod desc")

ord_prod %>%
  ggplot(
    aes(reorder(order_id, n_prod, function(x) x), 
        n_prod)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("order") +
  ylab("product") +
  theme_minimal()
```


* What is the most popular department? 
```{r}

```


## Whisker, magrittr
```{r}
name<-"Bartek"
paste0("My name is",name,".")
whisker
data<-list(name="Bartek")
whisker.render("My name is {{name}}.",data)
```

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

#* For given department, plot the most popular product (with templating department).
* For given department, give a list of 10 the most popular product (with templating department).
```{r}
product_times_ordered_txt <- "
SELECT op.product_id, product_name
,  COUNT(1) as n_orders
FROM (
        SELECT product_id
        FROM order_products__prior_tbl
    ) op
left JOIN (
        SELECT  product_name
        ,   product_id
        ,   department_id
        FROM products_tbl) p
ON op.product_id = p.product_id
WHERE department_id = {{departmentId}}
GROUP BY op.product_id, product_name
ORDER BY n_orders desc
"
data <- list(departmentId = 19)
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) 
```

* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
#
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name?

```{r}
"
SELECT op1.product_id
,  avg_products_number, p.product_name
From(select product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn 
GROUP BY product_id
ORDER BY avg_products_number DESC) op1 
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op1.product_id = p.product_id


" %>%
dbGetQuery(sc, .)
```

*Product that is usually orders in small baskets
```{r}
"
SELECT op1.product_id
,  avg_products_number, p.product_name
From(select product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn 
GROUP BY product_id
ORDER BY avg_products_number ) op1 
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op1.product_id = p.product_id


" %>%
dbGetQuery(sc, .)
```


*for 10 first users and all their orders find days since first order

```{r}
dbGetQuery(sc, 
"
SELECT  user_id
,   order_id
,   IF( first_order IS NULL, 0,  first_order) AS days_since_first_order
FROM
(SELECT user_id
,   order_id
, sum(days_since_prior_order) OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 ) aaa
")
```

